---
title: "Polycystic Ovary Syndrome (PCOS)"
output: pdf_document
---

# 1. Problem Statement

The goal of this analysis is to use the Polycystic Ovary Syndrome (PCOS) dataset from Kaggle (<https://www.kaggle.com/datasets/hasaanrana/diet-exercise-and-pcos-insights>) to determine if there are any factors that could be linked to having the disease.

# 2. Data Loading and Exploration

```{r message=FALSE, warning=FALSE}
library('dplyr')
library('ggplot2')
library('tidyr')
library('rmarkdown')
```

```{r}
pcos_data_raw <- read.csv('pcos_data.csv')
```

```{r}
dim(pcos_data_raw)
```
The dataset contains 173 rows and 36 columns.

```{r}
head(pcos_data_raw)
```
```{r}
colnames(pcos_data_raw)
```
```{r}
summary(pcos_data_raw)
```
```{r}
count(pcos_data_raw, PCOS, sort = TRUE)
```
The dataset contains 134 women without PCOS, 38 with PCOS, and one entry with both values.

# 3. Data Cleaning and Preparation

## 3.1 Dealing with Missing and Duplicated Values

## 3.2 Adding and Modifying Columns

# 4. Data Visualization in Tableau

# 5. Analysis

## 5.1 General Questions

## 5.2 Inferential Statistics

# 6. Conclusion
